---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(biomaRt)
library(GenomicRanges)
library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(plyr) 
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')


data = '~/Data/Ellie'
info.dir = '~/Data/Ellie/QDNAseq'
#tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

#includeDemo = F

#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = subset(all.patient.info, Set == 'Training')

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)


pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

all.info = rbind(sum.patient.data, sum.validation.info)

```

# Segment size: 5e06 with arms


## Quick look at our demographic information

```{r echo=F}

pander(
  as.data.frame(all.info %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range'=paste(range(end.year-start.year), collapse='-'),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='full cohort')
```


```{r echo=F}
cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

cohortSummary = cbind(cohortSummary,
  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
        table(subset(all.patient.info, Status == 'P')$Pathology)))


  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
        table(subset(patient.info, Status == 'P')$Pathology))
  
  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
        table(subset(validation.patient.info, Status == 'P')$Pathology))

```

`r pander(t(cohortSummary), caption='Full cohort, split by status')`


We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.


```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme


p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)

```

```{r, echo=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```


Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all. However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53. If we look at pathology the breakdown is even more obvious.  `r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.

`r pander(p53.pred.byPath, caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`


# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.


## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=F)
dim(dysplasia.df)
length(labels)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, coefs, performance.at.1se, models)

riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Global copy number plots

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
 
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

arm.loc = arrange(rbind(
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)


globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')


globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it 

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)

```

```{r loaddata, message=F, warning=F, echo=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}


cx = dysplasia.df[,'cx']

x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

```

# Building the model

We built the model using a 50-fold cross-validation process on the patients (rather than samples) for various lambda and alpha values to find the best fitting model. 

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

  do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

```

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```


## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

```


`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
`r pander(sapply(mp$stable.coeffients, length), caption='Total features at each value of alpha')`


`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors. 

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )

rm(performance.at.1se, coefs, mp)
```

#### LOO minus HGD

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```


```{r, echo=F, message=F, fig.width=5, fig.height=5}
pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)
  

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))
    
p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p
  
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)


rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
  

```


```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
  
  fitNOHGD = models[[select.alpha]]
  
  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']
  
  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[hgd.samples$Samplename,]
  
  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  hgd.samples$Prediction = pred
  
  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )
  
  rm(no.hgd.plots, coefs, performance.at.1se, models)
}

#hist(hgd.samples$Prediction)


rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```



### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```


```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
#classification.rate$x = 1:nrow(classification.rate)

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```


# Features

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

## Arms...
if (exists('arm.locs')) {
  ch[which(ch$label %in% with(arm.locs, paste(chr, ':', start, '-', end, sep=''))),]
}

```

`r pander(ch, caption='Selected coefficients')`


```{r nohgd_coefs, message=F, warning=F, echo=F}
setdiff(ch$label, rownames(nohgd.coefs))

```


```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
genomeFeat = grep('^\\d+:', ch$label)
featureAnn = data.frame(apply(do.call(rbind, strsplit(ch$label[genomeFeat], ':|-')), 2, as.numeric))
colnames(featureAnn) = c('chr','start','end')
featureAnn[] = lapply(featureAnn[], function(x) as.numeric(as.character(x)))

chh = cbind(ch[genomeFeat,], featureAnn)
chh = subset(chh, !is.na(chr))

chh = merge(chh, chr.lengths[,c('chrom','chr.length')], by.y='chrom', by.x='chr')
chh$chr = factor(chh$chr, levels=c(1:22))

chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

custom.max<-function(cf, hazards, ymax) {
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))

p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)


p = globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

```

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p

ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

#oddPt = names(pg.samp)[which.max(performance.at.1se)]
#subset(sum.patient.data, Hospital.Research.ID == oddPt)

#pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```


```{r, eval=F, echo=F, message=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

image(cor(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

chr <- as.numeric(sub(":.+","",colnames(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))])))
c <- which(diff(chr)==1)/length(chr)
abline(h=c, v=c, col='blue')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Samplename','Prediction')]))
rownames(preds) = preds$Samplename
p <- preds[names(labels), 'Prediction']

d <- dist(dysplasia.df[,c(1:(ncol(dysplasia.df)-3))])
t <- tsne::tsne(d)
plot(t,  bg=colorRampPalette(c('green','grey','red'))(10)[cut(p, breaks = seq(0,1,l=11))], pch=21, col=ifelse(labels==1,"black","white"))
```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

```

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'discovery'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)


insilico.rocs[['discovery']] = roc

df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
))

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)


df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
#    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme


roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
pROC::coords(rocNOHGD, "best")
pROC::coords(rocNOHGD, 0.6)
pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)


cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=8 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)
       
hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID)
)

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=4}

#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))

ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
 cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

ptendo = ptendo[-which(is.na(ptendo$mean)),]

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

```


### Compare ROCs


```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
  
  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))

  
  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}

get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
# To find the auc at a given spec or sens
  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
}

# get.auc.at(p53roc)
# get.auc.at(hgdpathroc)
# get.auc.at(pathroc)
# get.auc.at(roc)

roc$model = "CN Model"

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc,p53roc)

cols = brewer.pal(4, "PRGn")
dodge = position_dodge(width=0.9)

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

```


```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

m = rbind.data.frame(
      get.roc.stat(roc),
      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

cols = brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
```


## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
  
  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))
      
      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

  tb = table(pt[,c('months.before.final', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
    for (j in 2:length(yrs)) 
      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

back$Pathology = as.character(back$Pathology)
#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

ggplot(back, aes(Pathology, Prediction, group=Pathology, fill=Pathology)) + facet_grid(~Status) + 
  geom_jitter(width=0.2) + geom_boxplot() + plot.theme  + theme(legend.position = 'none')

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
 f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'


pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                      ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))


annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))
pcal

ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
```

```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
```

```{r indP, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
  geom_point(aes(color=Risk, shape=Status)) +scale_color_manual(values=riskCols, name='P(P)') + 
  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
  labs(x='sample', y='model probability', title='Per-sample probabilities')
```

```{r relrisk, echo=F, fig.width=6, fig.height=6}
#myPal = c(RColorBrewer::brewer.pal(9, 'Blues')[9:7], 
#          adjustcolor(grDevices::heat.colors(10)[6], blue.f=1.5),
#          ( grDevices::heat.colors(10)[7:6]),
#          RColorBrewer::brewer.pal(9, 'Reds')[6:9])

myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
    plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
predhist

ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)

#ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
```


### Mountain plots using spatial info

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
min.theme = theme(text=element_text(size=8), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   ) 


mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
  if (is.null(pal))
    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')

  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
  df$ogj = factor(df$ogj, ordered = T)
  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
  #pal = pal[which(limits %in% levels(df$nl))]

  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
  fills = c()
  for (yr in sort(unique(df$Endoscopy.Year))) {
    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
  }
    
  blanklabel <- function(s) {
    return('')
  }
  
  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
      labs(title=title, x='') + min.theme  
  
  legend = get.legend(p)
  p = p + theme(legend.position = 'none')
  
  g = ggplot_gtable(ggplot_build(p))
  strips = grep('strip-r', g$layout$name)

  for (n in 1:length(strips)) {
    i = strips[n]
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
  }

  #grid.arrange(g)
  return(list('grob'=g, 'legend'=legend))
  #grid::grid.draw(g)  
  #return(g)
}

back$ogj = back$nl
back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
back$ogj[which(back$nl >= 7)] = 3 

ogjLimits = levels(factor(back$ogj, ordered = T))

df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')

# early/late samples
#df = subset(globalCNdf, Patient == pt)
df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
df = arrange(df, chr, start, Endoscopy.Year)

## count arms differently than segs??
df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% group_by(Status, Patient) %>% mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))


el = subset(df2, early | late)
el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)




ggplot(el2, aes(Status,`diff(arms)`)) + geom_boxplot() + geom_jitter(width=0.2)
ggplot(el2, aes(Status,`diff(CNA)`)) + geom_boxplot() + geom_jitter(width=0.2)


p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )

# for (i in 1:nrow(sum.patient.data)) {
#   print(i)
#   pt = sum.patient.data[i, 'Patient']
#   status = sum.patient.data[i, 'Status']
# 
#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
#   
#   fl = paste('patient_', pt, '.png', sep='')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
# }

back$ogj = NULL

```


```{r, echo=F, fig.width=2, fig.height=4}

grid.arrange(p2$legend)

ggsave(filename='plots/spatial_mtn_pt20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
ggsave(filename='plots/spatial_mtn_pt51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)

```


At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`


## Jacknife coefficients

The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.

Could just as easily get a jacknife CI per patient

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )
  sharedcfs = intersect(names(lcf), ch$label)
  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

library(bootstrap)

# per patient
# jack.se = apply(coefs.per.pt, 2, function(x) {
#   jk = jackknife(x, var)
#   jk$jack.se
# })
# js = cbind.data.frame(performance.at.1se, jack.se)
# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() + 
#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))


# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')

```

Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.


## How do the predictions look at each time point?

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom', text=element_text(size=10))
rp

ggsave(filename='plots/pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)


```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}

#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=7, height=6, units='in', dpi=300)

```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
                                          'High'=sum(length(which(Risk == "High"))),
                                          'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)
ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) + 
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
  plot.theme

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology))

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
     breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),
  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

endos.m = arrange(merge(endos.m,tmp), brks)

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

`r pander(as.data.frame(pred.confidence))`

The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".


```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

#pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T, OR=F) + labs(title='') + theme(legend.position = 'none')
#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
```

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
ga = grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
ncol=2)
ga

ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)
  
  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

summary.t = table.path.risk(back)

write.table(summary.t, quote=F, sep='\t', row.names = F)
```


`r pander(as.data.frame(summary.t))`


```{r, echo=F, message=F, warning=F}
back = back %>% group_by(Patient) %>% mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')

ggsave('plots/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)

ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)

ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)

ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)


```


# Chr17

## p arm
```{r, echo=T}

p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
p11Loss = subset(p11, value < -1)
table(p11Loss$Status)

p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)

p11path['BE']/sum(p11path)
sum(p11path[c('HGD','IMC')])/sum(p11path)

table(p11Loss[c('Patient', 'Pathology')])

# Samples with a loss vs all Progressor samples
length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))

p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time

table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)

(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))


pi = patient.info
pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)

plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))


# So why might there be loss but no corresponding aberrant staining?
table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)

```

## Top Features
```{r, echo=T}
m = data.frame(matrix(nrow=nrow(chh), ncol=0))

for (i in 1:nrow(chh)) {
  feat = chh[i,]
  
  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
  if(feat$end-feat$start > 5e6)
    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')  
  
  segGL = subset(seg, value < -1)
  if(feat$coef > 0)
    segGL = subset(seg, value > 1)

  m[i, 'n.samples'] = length(unique(segGL$variable))
  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))
  
  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)

  m[i,'BE'] = segpath['BE']/sum(segpath)
  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
}
rownames(m) = chh$label

range(m$ratio.samples.P)
summary(m$ratio.samples.P)
subset(m, ratio.samples.P == max(ratio.samples.P))
subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
#table(cdkGain[c('Patient', 'Pathology')])

cor.test(chh$hazards, m$n.samples)

# Samples with a gain vs all samples

```

## Chr 17 Predictive

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

ccgenes = cbind(ccgenes, loc)
ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))

genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

ov = findOverlaps(chGR, genesGR)
#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
#colnames(genes)[1] = c('chr')


g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + labs(
title="All Samples") + theme(legend.position = 'none')
g17

g17 = g17 + theme(text=element_text(size=11), strip.text = element_text(size=11))
ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)

```


```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
chGR$Genes = ''
for (hit in unique(queryHits(ov))) {
  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
}

for (chr in unique(chh$chr)) {  
  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
  print(p)
}

annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2

p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5 ) + theme(legend.position = 'none')

ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)

```

`r pander(genes[,c(1:3,6,7)])`


```{r, echo=F, message=F, warning=F, fig.height=5, fig.width=8, eval=F}
file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("File missing", file))
}

names(performance.at.1se)

mp = model.performance(performance.at.1se, coefs) 
mp$plot = mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
  scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
  geom_hline(yintercept=mean(mp$performance$mean)+sd(mp$performance$mean), color='red', alpha=0.5) +
  geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)
mp$plot
ggsave('plots/performance/performance_subsets.png', plot=mp$plot, height=5, width=9, units='in',dpi=300)


chr.info = read.table('hg19_genes.txt', sep='\t', header=T)
chr.info = chr.info[1:22,]

chr.info$info.content = chr.info$protein_coding/chr.info$length

# No correlation between information content (genes/length) and classification rate 
cor.test(mp$performance[1:22,'mean'], chr.info$info.content)


median(mp$performance$mean)
sd(mp$performance$mean)
mp$performance[17,'mean']

```

Individually none of these show very good classification rates, all hovering around 65%. However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data

# Tumour Supressors

```{r ts, eval=F}
colnames(ccgenes)

tsgs = ccgenes
nrow(tsgs)

tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol')], keep.extra.columns = T)
tsgs = sort(tsgs)

length(tsgs)

# average 1 per 9MB
avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)


chh = arrange(chh, -hazards)
features = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

distanceTSG = data.frame(matrix(nrow=0,ncol=3))
for (i in 1:length(features)) {
  f = features[i]
  hits = nearest(f, tsgs, ignore.strand=T, select='all')
  if (length(hits) > 0) {
    distanceTSG = rbind(distanceTSG, cbind.data.frame(
      'features'=i,
      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
    )
  }
}

features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]

# In our dataset we should see 1 TSG/onc per 10MB
# But this is double the actual number we see
ncol(dysplasia.df)/round(avgTSGdist/5e6)

length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))
```


# Validation Dataset

20 Patients (10P, 10NP) set aside for validation purposes
`r pander(table(sum.validation.info$Status), caption='Validation cohort')`

```{r val_setup, echo=F, message=F, warning=F, include=F}
validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)

raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)

# Get the SD, mean, and variance(?) per sample
v.sampleVar = do.call(rbind, lapply(raw, function(f) {
  t = read.table(f, sep='\t', header=T)
  sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
  x = do.call(rbind, lapply(sampleCols, function(i){
    cbind('Sample.Mean'=mean(t[,i]),
          'Sample.SD'=sd(t[,i]),
          'Sample.Var'=var(t[,i]))
  }))
  rownames(x) = colnames(t)[sampleCols]
  return(x)
}))

arms = grep('arms', cleaned, value=T)
segs = grep('arms', cleaned, invert=T, value=T)

length(segs) == length(arms)

# Load segment files
tiled.segs = do.call(rbind, lapply(segs, function(f) {
  load.segment.matrix(f)
}))
dim(tiled.segs)

tiled.arms = do.call(rbind, lapply(arms, function(f) {
  load.segment.matrix(f)
}))

length(z.mean) == ncol(tiled.segs)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])

for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

dim(tiled.arms)
dim(tiled.segs)

cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)

validation.df = subtract.arms(tiled.segs, tiled.arms)
validation.df = cbind(validation.df, 'cx'=cx.score)

fauxlabel = rep(0, nrow(validation.df) )
names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))

```


```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])

locs = do.call(rbind.data.frame, lapply(rownames(valCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

valCNdf = cbind.data.frame(locs, valCNdf)
valCNdf = melt(valCNdf, id.vars=colnames(locs))
valCNdf$chr = factor(valCNdf$chr, levels=c(1:22), ordered = T)

valCNdf$seg.type = '5MB'
valCNdf[with(valCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
valCNdf = merge(valCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

valCNdf = merge(valCNdf, validation.patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
valCNdf$Status = ifelse(valCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

valCNdf$cn = factor(sapply(valCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
valCNdf$mean.value = valCNdf$value # but not going to use it 

p = cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) + theme(legend.position = 'none') + labs(title='Validation: All samples')
p
ggsave('plots/validation_mtn.png', plot=p, width=10, height=5, units='in', dpi=300)

```


```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
message(paste("loading", file))
load(file, verbose=T)

fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')

vpd = lapply(unique(validation.patient.info$Hospital.Research.ID), function(pt) {
  info = subset(validation.patient.info, Hospital.Research.ID == pt)
  info$Prediction = NA
  info$OR = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  
  info = cbind(info, v.sampleVar[info$Samplename,])
  
  return(info)
})
names(vpd) = unique(validation.patient.info$Hospital.Research.ID)


# for (s in rownames(pred)) {
#   vpd[which(vpd$Samplename == s), 'Prediction'] = pred[s,]
#   vpd[which(vpd$Samplename == s), 'OR'] = or[s,]
# }

for (pt in names(vpd)) {
  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
  vpd[[pt]]$OR[which(vpd[[pt]]$Samplename %in% rownames(or))] = or[ intersect(rownames(or), vpd[[pt]]$Samplename), ]
}

back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back.v$brks = with(back.v, factor(brks, rev(levels(brks))))

back.v$Pathology = as.character(back.v$Pathology)

back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

```


```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
back.v$Risk = factor(back.v$Risk, levels=c('High','Moderate','Low'))
back.v$nl = factor(back.v$nl, ordered = T)

sq = c(0, seq(1, max(back.v$months.before.final), 12))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, 
                 labels=c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-length(sq)],
                  include.lowest = T, ordered_result = T)
back.v$brks = with(back.v, factor(brks, levels=rev(levels(brks))))

m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
                                            'High'=length(which(Risk == 'High'))/length(Patient),
                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))

ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=riskCols) +
  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set') + plot.theme

back.v$Prediction

table(subset(back.v, Status == 'P')$Prediction > 0.6)
table(subset(back.v, Status == 'NP')$Prediction < 0.3)

```

```{r echo=F, warning=F, message=F, fig.height=10,fig.width=25}
back.v$Patient = factor(as.numeric(back.v$Patient), ordered=T)

p1 = pred.tiles(subset(back.v, Status == 'P'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Progressors', x='') + theme(legend.position = 'bottom') 
valLeg = get.legend(p1)
p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back.v, Status == 'NP'), cols=myPal, probs=T, nrow=1) + labs(title='Validation Non-Progressors') + theme(legend.position = 'none')

grid.arrange(p1, p2, nrow=2)

ggsave(filename='plots/val_heatmap.png', plot=grid.arrange(p1, p2, nrow=2), width=18, height=6, units='in', dpi=300)

ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)


```

```{r echo=F, warning=F, message=F, fig.height=5,fig.width=8}
vig.plot(subset(back.v, Patient == 68), subset(validation.patient.info, Patient == 68), cols=riskCols)
```


```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)

hgdimc = subset(riskPath, Pathology %in% c('HGD','IMC')) %>% group_by(Risk) %>% summarize (
  Freq=sum(Freq),  nPath=sum(nPath), ratio=sum(Freq)/sum(nPath))

riskPath = riskPath[-which(riskPath$Pathology %in% c('HGD','IMC')),] 
riskPath = rbind(riskPath , cbind('Pathology'='HGD/IMC', 'Status'='Progressor', hgdimc))

riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

pr = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) +   scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4, show.legend = F) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
pr

ggsave(filename='plots/val_riskpath.png', plot=pr, width=7, height=7, units='in', dpi=300)



```

## Validation prediction calibration

```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=3}
cuts = seq(0,1,0.1)
qt = as.data.frame.matrix(table(back.v$quants, back.v$Status))
qt$quant = rownames(qt)

qt = cbind(qt, back.v %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) )) 

pred.confidence.v = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 'perc'=P/sum(NP,P) )

pred.confidence.v[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence.v[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)


pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

ggplot(pred.confidence.v, aes(mn, perc)) + 
  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() +
  #geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + 
  labs(x='Mean prediction', y='Ratio of `P` patients', title='Validation calibration') +
  scale_color_manual(values=riskCols) + plot.theme 

preds = do.call(rbind.data.frame, lapply(vpd, function(df) df[c('Status','Prediction')]))
vroc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
vroc$model = 'Validation'

roc$model = 'Discovery'
multi.roc.plot(list(roc, vroc), title='ROC', palette='Dark2') + theme(text=element_text(size=12))
#ggsave('plots/auc_dv.png', plot=p, width=5, height=4, units='in', dpi=300, scale=2)
```


```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=4}
aucs = rbind.data.frame(pROC::ci.auc(roc), pROC::ci.auc(vroc))
colnames(aucs) = c('ci.min', 'auc','ci.max')
aucs$model = c('Discovery','Validation')

ggplot(aucs, aes(model, auc)) + geom_point(size=2) + ylim(0,1) +
  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + coord_flip() + 
  geom_text(aes(y=ci.min, label=round(ci.min, 2)), vjust=4) +
  geom_text(aes(y=ci.max, label=round(ci.max, 2)), vjust=-4) +
#  geom_text(aes(y=auc, label=round(auc, 2)), vjust=-1) +
  labs(x='', y='AUC', title='AUC with CI') + plot.theme 
```



## Downsampled Datasets

```{r echo=F, eval=T}
dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
rm(dysplasia.df,coefs,plots,labels)

select.alpha = '0.9'
fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']


predA = list.files(dirA, 'predictions.Rdata', full.names = T)
predB = list.files(dirB, 'predictions.Rdata', full.names = T)

files = c()
if (length(predA) < 1)
  files = c(list.files(dirA, pattern='raw', full.names=T))
if (length(predB) < 1)
  files = c(files, list.files(dirB, pattern='raw', full.names=T))

if (length(files) > 0) {
  sampleNames = sub('_raw.txt', '', basename(files))
  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
  
  for (n in 1:length(files)) {
    print(n)
    print(sampleNames[n])
    
    rs = read.table(files[n], sep='\t',header=T)
    if (ncol(rs) < 4)
      stop("File needs to contain 4 columns: chr, start, end, sample_value")
    
    segs <- tile.segmented.data(rs, size=5e6, chr.info=chrlen)
    segM = as.matrix(segs[,-c(1:3)])
    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
    segM = t(segM)
    for (i in 1:ncol(segM)) 
      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])
    
    arms <- tile.segmented.data(rs, size='arms', chr.info=chrlen)
    armsM = as.matrix(arms[,-c(1:3)])
    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
    armsM = t(armsM)
    for (i in 1:ncol(armsM)) 
      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])
    
    nrow(armsM) == nrow(segM)
      
    cx.score = score.cx(segM,1)
    
    mergedDf = subtract.arms(segM, armsM)
    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
    dim(mergedDf)
    
    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]
    
    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')
    
    dspred[n,] = c(preds[,1],RR[,1])
    print(dspred[1:n,])
  }
} else {
  load(predB, verbose=T)
  dspredB = dspred

  load(predA, verbose=T)
  
  dspred = rbind(dspred, dspredB)
  dspred$Type = NA
  
  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'
  
}

ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
  theme(legend.position = 'none') + plot.theme 

adj = subset(dspred, Type == 'AdjacentBE')
prg = back[,c('Prediction', 'OR', 'Status')]
prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')

#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
colnames(prg) = colnames(adj)
m = rbind(adj,prg)
m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T) 

pr = ggplot(m, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2) + geom_boxplot() + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
pr

ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)

#save(dspred, file=paste(dirB, 'predictions.Rdata', sep='/'))


```

# Clinical Story

First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.  Likely a consequence of sample size. The labels show the proprotions of progressors that fell into each quintile in the training set.

```{r, echo=F, warning=F, message=F}

summary.v = table.path.risk(back.v)

pander(as.data.frame(summary.v))

write.table(summary.v, sep='\t', quote=F, row.names = F)
```

## Vignettes

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')


## P
p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
vigLeg = get.legend(p)
p = p + theme(legend.position = 'none')
ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_86.png', plot=p, width=6, height=6,dpi=300)

print(p)

## NP
p = vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_3.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_23.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_33.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_49.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_51.png', plot=p, width=6, height=6,dpi=300)
           
p = vig.plot(subset(back, Patient == 6), subset(patient.info, Patient == 6),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_6.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 54), subset(patient.info, Patient == 54),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_54.png', plot=p, width=6, height=6,dpi=300)

print(p)

```

### Recommendations

 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions

```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}

rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    risks = table(Risk[i:(i+1)])
    p53 = table(p53.Status[i:(i+1)])
    rule = 'None'
    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
      rule = 1
    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
      rule = 2
    } else if ( risks['Moderate'] > 0 ) {
      rule = 3
    } else if ( risks['Low'] == 2) {
      rule = 4
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

cols = intersect(colnames(back), colnames(back.v))
#cols = cols[-grep('quants|pred\\.ratio|se\\(Risk\\)', cols)]
allB = rbind.data.frame(back[,cols], back.v[,cols])
allB$p53.Status = factor(allB$p53.Status, levels=c(0,1), ordered = T)

maxP = allB %>% group_by(Patient, Status, brks) %>% summarise( 'maxP'=max(Prediction), 'Risk'=Risk[Prediction == max(Prediction)], 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )

rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))

rules$t1 = factor(rules$t1, levels=levels(back$brks), ordered=T)
rules$t2 = factor(rules$t2, levels=levels(back$brks), ordered=T)

penult = subset(rules, t2 != 'Endpoint' & (t1 == rev(levels(back$brks))[2] | t2 == rev(levels(back$brks))[2] ))
round(table(subset(penult, Status == 'P')$rule)/nrow(subset(penult, Status == 'P')), 2)
round(table(subset(penult, Status == 'NP')$rule)/nrow(subset(penult, Status == 'NP')), 2)


second = subset(rules,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

table(subset(second, Status == 'P')$rule)
table(subset(second, Status == 'NP')$rule)


path.rx<-function(Patient, brks, Pathology) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    #risks = table(Pathology[i:(i+1)])
    rule = 'None'
    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
      rule = 1
    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
      rule = 2
    } else if ( length(which(grepl('BE|ID', Pathology[i:(i+1)]))) > 1  ) {
      rule = 3
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))
rules2$t1 = factor(rules2$t1, levels=levels(back$brks), ordered=T)
rules2$t2 = factor(rules2$t2, levels=levels(back$brks), ordered=T)

second = subset(rules2,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

table(subset(second, Status == 'P')$rule)
table(subset(second, Status == 'NP')$rule)



```


## Incidence

Median follow up for our discovery cohort was `r median(sum.patient.data$end.year-sum.patient.data$start.year)`. 

Adjust the RR of the orginal model with 1-3.5% offset to get an absolute risk (AR) with the understanding that the cohort was age matched and unbalanced.


```{r echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, coefs, cv.patient)

fit = models[[select.alpha]]
lambda = performance.at.1se[[select.alpha]]$lambda

# number of patients we'd needed to have included to get 45 progressors from a representative population (I think)
all = rbind(sum.patient.data, sum.validation.info)

# Per year
# This isn't quite right...we'd want to do something more like a 5 year 
#pop = round((table(all$Status)['P']/(0.07*100))*100)

```

```{r estimateRisk, echo=F, warning=F, message=F, eval=F}
# Probability that an individual will progress in a lifetime (20 years) given a rate of progression of 0.07 is about 77%, the probability for a specific individual is less
1-(0.93^20)

(1-(0.99^3))*15
(1-(0.99^3))/10
(1-(0.99^3))/10

# Or probability an individual will progress given a rate of 1%-3.5%
poft = sapply(seq(0.96,0.99, .01), function(x) {
 sapply(1:5, function(y) 1-(x^y) )
})

poft = as.data.frame(poft)
colnames(poft) = paste((1-seq(0.96,0.99, .01))*100, '%', sep='')

rp = apply(poft,1,function(x) cbind(mean(x),min(x), max(x)) )

rownames(rp) = c('mean','min','max')
colnames(rp) = paste('Year',1:5)

rr = range(as.data.frame(subset(back, Pathology %in% c('BE'), select=c('Patient','Pathology','months.before.final','OR')))$OR)

p = subset(back, Status == 'P' & months.before.final > 6 & OR > 10)$Patient
subset(sum.patient.data, Patient %in% p) %>% group_by(Patient) %>% summarise(final.endo-initial.endo)

rrPath = back %>% group_by(Status, Pathology) %>% summarise( 'mean.RR'=mean(OR),'ci.min'=mean(OR)-(1.96*sd(OR)), 'ci.max'=mean(OR)+(1.96*sd(OR)) )


rrp = ggplot(rrPath, aes(Status, mean.RR)) + #scale_color_gradientn(colors=myPal, limits=c(min(rrPath[,3:5]), max(rrPath[,3:5])),  name='') +
  facet_wrap(~Pathology, ncol=1) + 
  geom_point() +  geom_errorbar(aes(ymin=ci.min, ymax=ci.max), width=0.1) + 
  geom_text(aes(y=ci.min, label=round(ci.min,2)), vjust=2) + geom_text(aes(y=ci.max, label=round(ci.max,2)), vjust=2) +
  coord_flip() + plot.theme + labs(y='Mean RR', x='')
rrp

ggsave('plots/relrisk_path.png', plot=rrp, width=4, height=6,dpi=300)


as.data.frame(arrange(subset(back, Status == 'P' & months.before.final > 6 & Pathology %in% c('BE','ID','LGD') & OR > 1), Patient, months.before.final))

ggplot(t(rp), aes(1:5, mean)) +
  geom_ribbon(aes(ymin=mean/abs(rr[1]),ymax=mean*rr[2]), fill='rosybrown1') +
  geom_ribbon(aes(ymin=min,ymax=max), fill='grey88') +
  geom_line() + 
  geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=min, label='1%')) + geom_text(data=data.frame(t(rp[,5])), aes(x=5,y=max, label='4%')) +
  geom_line(aes(y=mean*rr[2]), color='red') +
  geom_line(aes(y=mean/abs(rr[1])), color='blue') +
  labs(y='P(P)', x='Year', title='Prob. of individual patient progression', subtitle='1-4% over 5 years') + plot.theme


subset(back, OR > 10)
cor.test(back$OR, as.integer(back$Pathology))
as.data.frame(subset(back, Pathology %in% c('HGD','IMC'), select=c('Patient','Pathology','months.before.final','OR')))

# as.data.frame(subset(back, Status == 'P' & OR < 2 & OR > -2, select=c('Status','Patient','Pathology','months.before.final','OR')))
# 
# hist(subset(back, Pathology %in% c('BE'))$OR)

```


```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
cases = table(sum.patient.data$Status)['P']

mn = round((table(sum.patient.data$Status)['P']/(0.01*100))*100)
m = round((table(sum.patient.data$Status)['P']/(0.0225*100))*100)
mx = round((table(sum.patient.data$Status)['P']/(0.035*100))*100)

offsetMin = log(cases/mn)
offsetMean = log(cases/m)
offsetMax = log(cases/mx)

#binom.test(cases, popL, 0.1)
# Offset for the linear predictor would then be -log(pop/44)
# Effectively this is my relative risk for the population
#offset = log(cases/popL)

# link function provides odds
#linPred = predict(fit, newx=dysplasia.df, s=lambda, type='link')
#quantile(linPred)

#df = rbind(dysplasia.df, validation.df)
#df = dysplasia.df

#p0 = predict(fit, newx=df, s=lambda, type='response')
#p0 = 1/(1+exp(-linPred)) # logit to get probabilities, same as type='response'
#p1 = 1/(1+exp(-linPred+abs(offset))) # calculates probs from risk (offset)

cols = c('Hospital.Research.ID','Patient','Status','PID', 'Samplename','Endoscopy.Year','Pathology','Prediction','OR')
#loopreds = rbind(loopreds[cols], do.call(rbind, vpd)[cols])
loopreds = do.call(rbind, pg.samp)[cols]

loopreds$P0 = 1/(1+exp(-loopreds$OR))

loopreds$Mean = 1/(1+exp(-loopreds$OR+abs(offsetMean))) # Adjust OR for RR
loopreds$Min = 1/(1+exp(-loopreds$OR+abs(offsetMin))) # Adjust OR for RR
loopreds$Max = 1/(1+exp(-loopreds$OR+abs(offsetMax))) # Adjust OR for RR

#ql = cbind.data.frame('y'=quantile(loopreds$OR)[c(2,4)], 'labels'=paste(c(25,75), '%', sep=''))

# From my nested-LOO predictions

ggplot(melt(loopreds[c('Min','Mean','Max')], measure.vars='Mean'), aes(value, fill=variable)) + 
  geom_density(alpha=0.5) + plot.theme


p = ggplot(melt(loopreds[c('Min','Max')]), aes(value)) + facet_wrap(~variable, ncol=2) +
  geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
    scale_fill_gradientn(colors = myPal,  name='') + 
    plot.theme + theme(legend.position = 'none') + 
    labs(y='n Samples', x='Absolute Risk',
         title='5-year adjusted risk', subtitle='Adjusted for upper & lower bound: 1-3.5% risk')
p

ggsave('plots/adj_RR.png', plot=p, width=6, height=6,dpi=300)


# ggplot(loopreds, aes(Min)) +
#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
#     scale_fill_gradientn(colors = myPal,  name='') + 
#     plot.theme + theme(legend.position = 'none') + 
#     labs(y='n Samples', x='Absolute Risk',
#          title='Sample Predictions: 5 year adjusted risk', subtitle='Adjusted for lower bound, 1% risk')
# 
# ggplot(loopreds, aes(Max)) +
#     geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + 
#     scale_fill_gradientn(colors = myPal,  name='') + 
#     plot.theme + theme(legend.position = 'none') + 
#     labs(y='n Samples', x='Absolute Risk',
#          title='5-year adjusted risk', subtitle='Adjusted for upper bound, 3.5% risk')

```


```{r lifetimeFPR, echo=F, fig.height=6, fig.width=6, eval=F}
## FPR estimates based on lifetime risks
# We could consider any NP patient with a minimum of 2 High risk calls?  or High/Moderate Risk?
# More correctly it should be any patients with concurrent High risk, but that's been pretty stable so two high is good enough for now

bd = bind_rows(back, back.v)
table(bd$Status)

risksL = subset(bd, Status == 'NP') %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

risksL = subset(risksL, high.risk > 0 | mod.risk > 0)

high = length(subset(risksL, high.risk >= 2 )$Patient)
#mod = length(subset(risksL, mod.risk >= 2 )$Patient)

# How would I tune it so that the FPR is a lot closer to the real incidence ~1%
# Current false positive rate is 1.5x higher than the highest incidence ~0.7
fpr=high*10/44
x = subset(bd, Status == 'NP' & Patient %in% subset(risksL, high.risk >= 2)$Patient & Risk == 'High')
(length(unique(subset(x, Prediction > 0.8)$Patient))*10)/44

# How would that affect TPR?
#pROC::coords(roc, 0.6)
#pROC::coords(roc, 0.8)

## Per year risk
## Evaluate over 5 years? How?
pyr = c()
for (yr in names(which(table(bd$Endoscopy.Year) > 10))) {
  risks = subset(bd, Status == 'NP' & Endoscopy.Year == yr) %>% group_by(Patient) %>% summarise( 'n.samp'=length(Patient), 'high.risk'=length(which(Risk == 'High')), 'mod.risk'=length(which(Risk == 'Moderate')) )

  risks = subset(risks, high.risk > 0 | mod.risk > 0)
  high = length(subset(risks, high.risk >= 2 )$Patient)
  mod = length(subset(risks, mod.risk >= 2 )$Patient)

  print(paste(yr,high*15, sep=':'))
  pyr = c(pyr, high*15)
}

summary(pyr)

pop*.07
pop*.02
pop*.05


```



```{r clonality, echo=F, message=F, warning=F, eval=F}
## Clonal Heterogenity
patient.info %>% group_by (Hospital.Research.ID, Patient, PID) %>% summarise(
  length(PID)
)

arms = c(grep(with(chr.lengths, paste(paste(chrom, ':', '1-\\d+000', sep=''), collapse='|')), colnames(dysplasia.df)),
         grep(with(chr.lengths, paste(paste(chrom, ':', '\\d+000-', chr.length, sep=''), collapse='|')), colnames(dysplasia.df)))


x = do.call(rbind, lapply(unique(patient.info$PID), function(pid) {
    df = as.matrix(dysplasia.df[subset(info, PID == pid)$Samplename, -grep('cx',colnames(dysplasia.df))])
    if (ncol(df) > 1) {
      cbind('arms'=sum(rowSums(df[,arms],na.rm=T))/nrow(df),
            'seg'=sum(rowSums(df[,-arms],na.rm=T))/nrow(df))
    } else {
      cbind('arms'=sum(df[arms,],na.rm=T),
            'seg'=sum(df[-arms,],na.rm=T))
    }
}))
rownames(x) = unique(patient.info$PID)
x = x[complete.cases(x),]

hist(x[,'seg'])
hist(x[,'arms'])

p = pnorm(x[,'seg'], mean(x[,'seg']), sd(x[,'seg']))
summary(p)

pa = pnorm(x[,'arms'], mean(x[,'arms']), sd(x[,'arms']))
summary(pa)

x[,'arms'] = scale(x[,'arms'])
x[,'seg'] = scale(x[,'seg'])



plot(x[unique(subset(patient.info, Patient == 48)$PID), 'arms'])
plot(x[unique(subset(patient.info, Patient == 48)$PID), 'seg'])

plot(x[unique(subset(patient.info, Patient == 34)$PID), 'arms'])
plot(x[unique(subset(patient.info, Patient == 34)$PID), 'seg'])

pp = back %>% group_by(Patient, Hospital.Research.ID, PID) %>% summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction)
)

pp = merge(pp, x, by.x='PID', by.y='row.names')

cor.test(pp$seg, pp$max)



```


## Cox model - Time to progress?

```{r cox, warning=F, message=F, echo=F, eval=T}
library(survival)

cox.pvalue<-function(cox, test='logrank') {
  pvalue = NULL  
  csum = summary(cox)
  
  if (test == 'logrank') {
    pvalue = csum$sctest[['pvalue']]  
  } else if (test == 'wald') {
    pvalue = csum$waldtest[['pvalue']]
  } else if (test == 'loglik') {
    pvalue = csum$logtest[['pvalue']]
  }
  return(pvalue)
}

prog.pred = lapply(pg.samp, function(df) {
  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'Age.at.diagnosis', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
  df = arrange(df, Endoscopy.Year, Pathology)

  df$Dup.PID = duplicated(df$PID)
  x = df[!df$Dup.PID,]
  
  for (i in 1:nrow(x)) {
    if (i == 1) {
      x[i,'tstart'] = 0
      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
      next
    }  

    x[i,'tstart'] = x[(i-1), 'tstop']

    # Final sample is different -- will leave it out later
    if (i == nrow(x)) {
      x[i, 'tstop'] = x[i,'tstart'] + 0.5
      break
    }

    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']

    # Samples taken in the same year
    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
      x[i,'tstop'] = x[i,'tstop'] + 0.5 
  }

  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
  
  if (sum(x$Diagnosis) >= 1)
    x = x[-which(x$Diagnosis == 1),]

  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
    x[nrow(x), 'Diagnosis'] = 1

  if (nrow(x) > 0) {
    x$Dup.Samples = ''
    for (j in 1:nrow(df[df$Dup.PID,]))  {
      row = df[df$Dup.PID,][j,]
      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
    }
    x$Dup.Samples = sub('^,','',x$Dup.Samples)
  }

  return(x)
})
## By pathology?
pp = do.call(rbind, prog.pred)


pp[is.na(pp$tstop), 'tstop'] = 0.5
pp$Pathology = as.character(pp$Pathology)
pp$AAD = cut(pp$Age.at.diagnosis, breaks = 3)

pp = merge(pp, back[c('Samplename','Risk', 'quants')], by='Samplename')
pp$Risk = as.character(pp$Risk)

sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Risk, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Risk')

cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Risk, data=pp) 
ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ Risk" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
#summary(cox)

sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ AAD, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Age') 
cox = coxph( Surv(tstart, tstop, Diagnosis) ~ AAD, data=pp) 
ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ Age at diagnosis" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
#summary(cox)


sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F) + xlim(0,5) + labs(y='Diagnosed', title='~Pathology')
cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples')) + xlim(0,5)
#summary(cox)


## All genomic regions
sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
sample.pred$Pathology = as.character(sample.pred$Pathology)

gcols = grep('\\d+:', colnames(sample.pred), value=T)
sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology','Age.at.diagnosis',gcols)]

sample.pred$Age.bracket = as.character(cut(sample.pred$Age.at.diagnosis, breaks=3))

# all
#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
#ggsurv(survfit(cox))
#summary(cox)
```
### Cox GLM selected features

```{r echo=F, warning=F, echo=F}
# glm selected features + Path
cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  grep('^\\d+:', ch$label, value=T))]) 
ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ GLM features + Pathology" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
#summary(cox2)


cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis','Age.bracket', grep('^\\d+:', ch$label, value=T))]) 
ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ GLM features + Age" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
#summary(cox2)

cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis',grep('^\\d+:', ch$label, value=T))]) 
ggsurv(survfit(cox2),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ GLM features" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox2),2),"\n", nrow(pp), 'samples')) 
#summary(cox2)


# chr arms only
options(scipen=999) # turn sci notation off
coxArms = coxph(Surv(tstart, tstop, Diagnosis) ~., data=sample.pred[,c('tstart', 'tstop','Diagnosis', with(arm.loc, paste(chrom, ':', start, '-', end, sep='')))])
ggsurv(survfit(coxArms),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ chr arms" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(coxArms),2),"\n", nrow(pp), 'samples')) 
#summary(coxArms)

# chr 17 only
cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox17),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ chr17" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox17),2),"\n", nrow(pp), 'samples')) 
#summary(cox17)

cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox11))
#summary(cox11)

cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox13))
#summary(cox13)

```


## CoxRFX

```{r echo=F, message=F, warning=F, eval=T}
plot.waldtest<-function(wt) {
  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
  if (length(grep('hazards', colnames(wt))) > 0) {
      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
        geom_point(aes(coef.x, coef.y, color=p.value<0.05, size=hazards), show.legend = F) +
        #scale_shape_manual(values=c(17,18), na.value=21) +
        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
        #geom_text_repel( aes( label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
  } else {
    p = p + geom_point(alpha=0.5) + 
      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
  }
  return(p + labs(title='Wald test'))
}
```

```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}

library(CoxHD)

coxDF = sample.pred

pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
  cbind( as.integer(coxDF$Pathology == p))
}))
colnames(pathology) = unique(coxDF$Pathology)


# age = do.call(cbind.data.frame, lapply(unique(coxDF$Age.at.diagnosis), function(p) {
#   cbind( as.integer(coxDF$Age.at.diagnosis == p))
# }))
# colnames(age) = unique(coxDF$Age.at.diagnosis)
age = coxDF$Age.at.diagnosis


## GLM regions only
coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], age, coxDF[,grep('^\\d+:', ch$label, value=T)])
coxDF$Pathology = NULL

gcols = colnames(coxDF)
cols = grep('^\\d+', colnames(coxDF))
colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')

coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))

#surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
surv = with(coxSM, Surv(tstart, tstop, Diagnosis))
groups = factor(c(rep('clinical',3),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
rfx = CoxRFX(coxSM[,-c(1:3)], surv, groups, verbose=T, nu=1)
summary(rfx)

plot(rfx)

wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
wt = merge(wt, ch[,c('label','hazards', 'coef','stability')], by='label', all=T) 

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')


options(scipen = 999)
# arms only
arms = with(arm.loc, paste(chrom, ':', start, '-', end, sep=''))
coxDfArms = coxDF[,arms]
colnames(coxDfArms) = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDfArms), value=T)), sep='') 

surv = with(coxDF, Surv(tstart, tstop, Diagnosis))
groups = factor(c(rep('clinical',3), rep('genomic', ncol(coxDfArms))))

rfx = CoxRFX( cbind.data.frame(coxSM[,c('ID','BE','age_10')], coxDfArms), surv, groups, verbose=T, nu=1)
summary(rfx)

plot(rfx)
wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))


```



```{r, eval=F}
# TODO: Relative risk
pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)
?predict
?predict.glm
quantile(pred)
p0 = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda, type='response')
p0
p1 = 1/(1+exp(-pred+2.66))
plot(p0,p1)
hist(p1)
hist(p0)
hist(p1)
plot(p0,p1)
hist(p1, n=20)
mean(p1>.5)
mean(p0>.5)
hist(pred)

```









