---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(biomaRt)
library(GenomicRanges)
library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(plyr) 
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')



data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

includeDemo = F

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = subset(all.patient.info, Set == 'Training')

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

validation.patient.data = patient.data[sum.validation.info$Hospital.Research.ID]
length(validation.patient.data)

patient.data = patient.data[sum.patient.data$Hospital.Research.ID]
length(patient.data)

patient.data = patient.data[!is.na(names(patient.data))]
names(patient.data)

sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## one patient was removed from the study
nrow(sum.patient.data)

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

all.info = rbind(sum.patient.data, sum.validation.info)

```

# Segment size: `r as.character(tile.w)` with demographics: `r includeDemo`


```{r echo=F}

pander( 
  as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='Full cohort')

```


## Quick look at our demographic information

We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.


```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme


p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)

```

```{r, echo=F}
x = do.call(rbind.data.frame, lapply(patient.data, function(df) df$info[,c('Status','p53.Status')]))
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind( 'Non-progressor'=summary(subset(x, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(x, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```


Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all. However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53. If we look at pathology the breakdown is even more obvious.  `r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.

`r pander(p53.pred.byPath, caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`


# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.


## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = plyr::arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)


cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, coefs, performance.at.1se, models, cv.patient)

```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Global copy number plots

```{r echo=F, warning=F, message=F, eval=T, fig.width=12, fig.height=8}

chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

arm.loc = arrange(rbind(
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% group_by(chrom) %>% mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)


globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')


globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it 


globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
globalCN.plot = globalCN.plot + theme(legend.position = 'none')

nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
#globalCNBEdf = subset(globalCNdf, Pathology == 'BE') %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

globalCN.plot 
globalBECN.plot


dfpts = subset(globalCNdf, Patient %in% c('57','76') & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

pt.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtCN.plot = cn.mtn.plot(dfpts, labeller(Status = pt.status))

dfpts = subset(globalCNdf, Patient %in% c('57','76') & Pathology == 'BE' & seg.type != 'arms')
#dfpts = dfpts %>% group_by(Patient, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )

ptbe.status <- c('Non-Progressor'=paste('Patient 57 (n=',length(unique(subset(dfpts, Patient == 57)$variable)), ')', sep=''),
              'Progressor'=paste('Patient 76 (n=',length(unique(subset(dfpts, Patient == 76)$variable)),')', sep=''))
genomePtBECN.plot = cn.mtn.plot(dfpts,  labeller(Status = ptbe.status )) + labs(title='NDBE samples only')

genomePtCN.plot 
genomePtBECN.plot

#cn.mtn.plot(dfpts,  labeller(Status = ptbe.status ), type='bar') + labs(title='NDBE samples only')

a = sample(subset(patient.info, Status == 'NP' )$Samplename, 1)
b = sample(subset(patient.info, Status == 'P' )$Samplename, 1)

cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both)) + labs(title="Individual samples")
cn.mtn.plot(subset(globalCNdf, variable %in% c(a,b) & seg.type != 'arms'), labeller(label_both), type='bar') + labs(title="Individual samples")

```



# Building the model

We built the model using a 50-fold cross-validation process on the patients (rather than samples) for various lambda and alpha values to find the best fitting model. 

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}

classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```


`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
`r pander(sapply(mp$stable.coeffients, length), caption='Total features at each value of alpha')`


`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors. 

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]])
                              )
```


### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

```


```{r, echo=F, fig.height=5, fig.width=5}
mp = model.performance(performance.at.1se, coefs)
mp$plot + labs(title=paste('No LGD/HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F,fig.height=5, fig.width=5}
#classification.rate$x = 1:nrow(classification.rate)

ggplot(classification.rate, aes(model, mean, group=model, color=model)) + 
  geom_text( aes(label=round(mean, 3)), hjust=-0.5, size=5, color='gray39', show.legend = F) +
  geom_text( aes(y=mean+sme*2, label=paste(all.coef, "\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  geom_point(size=4) +
  labs(x='', y='Classificaiton rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none')

```


# Features

```{r, echo=F, warning=F, message=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 


## Arms...
if (exists('arm.locs')) {
  ch[which(ch$label %in% with(arm.locs, paste(chr, ':', start, '-', end, sep=''))),]
}


write.table(ch, quote=F, sep='\t', row.names = F, col.names = T, file=paste( paste(tile.w, collapse='_')  ,'_features.txt'))
```

`r pander(ch, caption='Selected coefficients')`


```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
genomeFeat = grep('^\\d+:', ch$label)
featureAnn = data.frame(apply(do.call(rbind, strsplit(ch$label[genomeFeat], ':|-')), 2, as.numeric))
colnames(featureAnn) = c('chr','start','end')
featureAnn[] = lapply(featureAnn[], function(x) as.numeric(as.character(x)))

chh = cbind(ch[genomeFeat,], featureAnn)
chh = subset(chh, !is.na(chr))

chh = merge(chh, chr.lengths[,c('chrom','chr.length')], by.y='chrom', by.x='chr')
chh$chr = factor(chh$chr, levels=c(1:22))

chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

custom.max<-function(cf, hazards, ymax) {
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))

globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

ch$ymax = ifelse(ch$coef > 0, max(genomePtCN.plot$data$value), min(genomePtCN.plot$data$value))
chh = chh %>% rowwise() %>% mutate('max'=custom.max(coef, hazards, ymax))
genomePtCN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

genomePtBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F)

```


```{r echo=F, fig.height=5, fig.width=6}
globalCNdf$g.start = globalCNdf$start
globalCNdf$g.end = globalCNdf$end
for (i in 2:nrow(chr.lengths)) {
  rows = which(globalCNdf$chr == chr.lengths[i, 'chrom'])
  globalCNdf[rows, 'g.start'] = globalCNdf[rows, 'start'] + chr.lengths[i-1, 'genome.length']
  globalCNdf[rows, 'g.end'] = globalCNdf[rows, 'end'] + chr.lengths[i-1, 'genome.length']
}


globalCNdf = arrange(globalCNdf, chr, start)
globalCNdf$segment = with(globalCNdf, paste(chr, ':', start, '-', end, sep=''))

globalCNdf$chr = factor(globalCNdf$chr, levels=c(1:22), ordered = T)
globalCNdf$Status = factor(globalCNdf$Status)

p1 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(segment, variable, fill=value, group=Status)) + geom_tile() +
  scale_fill_gradient2(high='red3', low='green4', mid='white')  + 
  facet_grid(~chr, scales = 'free_x', space = 'free') + plot.theme + labs(x='',y='Samples') + 
  theme(axis.ticks=element_blank(), axis.text=element_blank(), panel.spacing.x = unit(0, 'lines'), legend.position='bottom')

p1leg = get.legend(p1)
p1 = p1 + theme(legend.position = 'none')

p2 = ggplot(subset(globalCNdf, seg.type != 'arms'), aes(y=variable, group=Status)) + 
  geom_tile(aes(x=0, fill=Status), show.legend = F) + 
  labs(x=' ', y='') + scale_fill_manual(values=c('darkblue','darkred')) + scale_x_continuous(expand=c(0,0)) +
  plot.theme + theme( axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid=element_blank())

grid.arrange(p1, p2, p1leg, ncol=2, nrow=2, widths=c(20,1), heights=c(10,1))
  

```

```{r echo=F, fig.height=4, fig.width=8}
ggplot(chh, aes(x=label,y=coef,fill=gl)) + geom_bar(stat='identity') + labs(x='Effects', y='') +
  coord_flip() + scale_fill_manual(values=c('darkgreen','darkred')) +
  plot.theme + theme(axis.line.x=element_blank(), axis.ticks=element_blank(),axis.text=element_blank(), panel.grid=element_blank(),panel.grid.major=element_blank(), legend.position='none')
```

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = coefs

df = as.data.frame(performance.at.1se)
ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='performance at lambda.1se', x='', title=paste('Performance for nested leave one out, at alpha=',select.alpha,sep='')) +
  plot.theme


#oddPt = names(pg.samp)[which.max(performance.at.1se)]
#subset(sum.patient.data, Hospital.Research.ID == oddPt)

#pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```


```{r, eval=F, echo=F, message=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

image(cor(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

chr <- as.numeric(sub(":.+","",colnames(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))])))
c <- which(diff(chr)==1)/length(chr)
abline(h=c, v=c, col='blue')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Samplename','Prediction')]))
rownames(preds) = preds$Samplename
p <- preds[names(labels), 'Prediction']

d <- dist(dysplasia.df[,c(1:(ncol(dysplasia.df)-3))])
t <- tsne::tsne(d)
plot(t,  bg=colorRampPalette(c('green','grey','red'))(10)[cut(p, breaks = seq(0,1,l=11))], pch=21, col=ifelse(labels==1,"black","white"))
```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=4, fig.width=4}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), c(rep(1, length(preds)), rep(0, length(np.preds))), main='Progressors & Non-progressors, cutoff 0.5')
roc50$model = 'CN Model 0.5'

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')
roc$model = 'CN Model'

grid.arrange(roc.plot(roc50,'Probability threshold = 0.5'), roc.plot(roc, 'Probability threshold calculated by ROC'), ncol=2)

roc.plot(roc, 'Copy number model')

cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=4 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]

pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'Dysplasia (HGD & LGD)'

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Compare ROCs

```{r echo=F, fig.height=6, fig.width=8}
get.roc.stat<-function(roc) {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
  
  sens = c(pROC::coords(roc, 'best')[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, 'best')[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, 'best')[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, 'best')[['sensitivity']], conf.level=0.95))

  
  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','Dysplasia (HGD & LGD)','CN Model'), ordered = T)


rocList = list(roc,roc50, pathroc,hgdpathroc,p53roc)

aucs = do.call(rbind, lapply(rocList,function(r) cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))

df = do.call(rbind, lapply(rocList, function(r) 
              cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model) 
              ))

ggplot(df, aes(Specificity, Sensitivity,color=model,fill=model)) + geom_line(show.legend = F) + scale_x_reverse() + 
    geom_label_repel(data=aucs, aes(label=paste(model,'\nAUC:', sprintf("%.2f",AUC), '%', sep='')), nudge_x=.25, color='white', show.legend = F) + 
  labs(title='ROC Curves', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme


dodge = position_dodge(width=0.9)
ggplot(melt(m, measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(label=paste(round(value,2)*100,'%',sep=''), y=0, vjust=-1 ), position=dodge, color='white')  +
  scale_fill_manual(values=c('purple3','dodgerblue3','firebrick3'), name='') + plot.theme + 
  theme(legend.position = 'bottom') + labs(x='', y='Percent', title='ROC Comparison') 
```

*** 

__All predictions from here onwards are performed using the `r cutoff` cutoff__

***

## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12
  )

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)
  
  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
      tmp = pt[which(pt$Endoscopy.Year == yr),]
      tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))
      
      match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
      pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

  tb = table(pt[,c('months.before.final', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
    for (j in 2:length(yrs)) 
      pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

back$Pathology = as.character(back$Pathology)
back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)
#cuts = quantile(back$Prediction, probs=seq(0,1,1/5))

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
 f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction), 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))

ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')


pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )
pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
    pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'


annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

#annotation.table = as.data.frame(annotation.table)
#rownames(annotation.table) = annotation.table$Risk
#annotation.table$Risk = NULL

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = c('green3','orange2','red3'), col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )

riskCols = c('red3','orange','green3')
grid.arrange(
  ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
    scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
    #geom_vline(xintercept=pred.confidence$r2, linetype='longdash', color='grey39') +
    #geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=75,label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + scale_color_manual(values=c('grey','black')) + 
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% group_by(Risk, min,max) %>% summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') 
  
  ,tableGrob(annotation.table, rows=NULL, theme=tt3),
  nrow=2, heights=c(8,2)
  )

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

ggplot(back, aes(1:nrow(back), Prediction, color=Risk)) + 
  geom_point(aes(color=Risk, shape=Status)) + geom_errorbar(aes(ymin=Prediction-`se(Risk)`, ymax=Prediction+`se(Risk)`)) +
  geom_hline(data=annotation.table, aes(yintercept=max), color='grey') + plot.theme +
  labs(x='sample', y='model probability', title='Per-sample probabilities')
```

At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`


## Jacknife coefficients

The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.

Could just as easily get a jacknife CI per patient

```{r jacknife, echo=T, warning=F, message=F, fig.height=5, fig.width=6}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
  lcf = loo.coefs[[pt]][,1]
  names(lcf) = rownames( loo.coefs[[pt]] )
  sharedcfs = intersect(names(lcf), ch$label)
  coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

library(bootstrap)

# per patient
# jack.se = apply(coefs.per.pt, 2, function(x) {
#   jk = jackknife(x, var)
#   jk$jack.se
# })
# js = cbind.data.frame(performance.at.1se, jack.se)
# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() + 
#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))


# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
  jk = jackknife(x, var)
  jk$jack.se
})
ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + 
  coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')

```

Effectively it looks to me like my coefficients have fairly tight CI if the SE of the jackknife variance is used.


## How do the predictions look at each time point?

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=9}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red2','orange','green3')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')


ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=Risk)) + geom_bar(stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red2','green3')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

```


```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% summarise( 'Low'=sum(length(which(Risk == "Low"))),
                                          'Moderate'=sum(length(which(Risk == "Moderate"))),
                                          'High'=sum(length(which(Risk == "High"))),
                                          'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)
ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=c('green3','orange2','red')) + 
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
  plot.theme

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% group_by(Status, months.before.final, brks, PID) %>% summarise(
              'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology))

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
     breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),
  'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
     labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% group_by(brks) %>% summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

endos.m = arrange(merge(endos.m,tmp), brks)

ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1)) 

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```


```{r, echo=F, fig.height=8, fig.width=6, eval=F}
## Progressor predictions per patient
#plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
```

```{r, echo=F, fig.height=8, fig.width=6, eval=F}
# plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
# plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
#plot.endo( subset(endos, Patient %in% subset(me, months > -48)$Patient), bin=T) + labs(title="Progressors", subtitle=paste(1, "-",48/12,  "years or more follow-up"))
```


```{r, echo=F, fig.height=12, fig.width=8, eval=F}
## Non-progressor predictions per patient
# nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)
# 
# nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
#   df = do.call(rbind, lapply(nonp.back, function(x) {
#     if (i > nrow(x)) return(NA)
#     cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
#   }))
#   rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient
# 
#   df$Patient = rownames(df)
# 
#   return(df)
# }))
# nonp.endos = nonp.endos[complete.cases(nonp.endos),]
# #nonp.endos$i = nonp.endos$i-1
# 
# nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
# nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
# nonp.endos$months.before.final = nonp.endos$months.before.final*-1
# 
# me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

#plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + labs(title='Non-progressors with progressive predictions')
```

`r pander(as.data.frame(pred.confidence))`

The first two quintiles and the last two together account for 80% of our samples.  Only 20% fall into the category "Moderate".


```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=10 }
back$nl = factor(back$nl, ordered=T)
back$labels = factor(back$Risk, levels=c('High','Moderate','Low'))

 pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), nrow=1) + labs(title='Progressors')

grid.arrange( pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1) + labs(title='Progressors') + theme(legend.position = 'none'),
              pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1) + labs(title='Non-Progressors'), ncol=2 )

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)
  
  summary.t = summary.t %>% group_by(Pathology) %>% mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$labels))]
}

summary.t = table.path.risk(back)

write.table(summary.t, quote=F, sep='\t', row.names = F)
```


`r pander(as.data.frame(summary.t))`


```{r, echo=F}
    
p = pred.tiles(subset(back, Status == 'NP'), nrow=2) + labs(title='Non-Progressors')
ggsave('npg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)
 

p = pred.tiles(subset(back, Status == 'P'), nrow=2) + labs(title='Progressors')
ggsave('pg_heatmap.png', plot = p, scale = 1, width = 52, height = 8, units = "in", limitsize = F, dpi = 300)

```


# Chr17

## p arm
```{r, echo=F}

p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')

p11Loss = subset(p11, value < -1)
table(p11Loss$Status)

p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)

p11path['BE']/sum(p11path)
sum(p11path[c('HGD','IMC')])/sum(p11path)

# Samples with a loss vs all samples
length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))


p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time

table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)

pander(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))

```

## Chr 17 Predictive

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]
ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations

genes = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & !grepl('^X|Y', Chr.Band), select=c('Gene.Symbol','Genome.Location'))


loc = unlist(strsplit(genes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(genes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

genes = cbind(genes, loc, 'value'=min(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))



plot.cn.chr(subset(globalCNdf, seg.type != 'arms'), chrom='17', chr.lengths, genes=genes, haz=chh) + labs(
title="Segment Values Samples", subtitle=paste('NonP:Prog - ', paste(table(patient.info$Status), collapse=':')))
```


```{r echo=F, warning=F, message=F}

plot.cn.chr(globalCNdf, chrom='8', chr.lengths, genes=genes, haz=chh)

plot.cn.chr(globalCNdf, chrom='3', chr.lengths, genes=genes, haz=chh)

plot.cn.chr(globalCNdf, chrom='21', chr.lengths, haz=chh, genes=genes)

plot.cn.chr(globalCNdf, chrom='11', chr.lengths, haz=chh, genes=genes)

plot.cn.chr(globalCNdf, chrom='16', chr.lengths, haz=chh, genes=genes)

plot.cn.chr(globalCNdf, chrom='9', chr.lengths, haz=chh, genes=genes)

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

ov = findOverlaps(chGR, genesGR)

#as.data.frame(chGR[unique(queryHits(ov))])
#length(genesGR[unique(subjectHits(ov))]$Gene.Symbol)
write.table(genesGR[unique(subjectHits(ov))]$Gene.Symbol, quote=F, row.names=F, sep='\t')

pander(as.data.frame(genesGR[unique(subjectHits(ov))])[,c(1:3,6,7)])

```


```{r, echo=F, message=F, warning=F, fig.height=5, fig.width=8, eval=F}
file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("File missing", file))
}

names(performance.at.1se)

mp = model.performance(performance.at.1se, coefs) 
mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
  scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
  geom_hline(yintercept=mean(mp$performance$mean)+sd(mp$performance$mean), color='red', alpha=0.5) +
  geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)

mp$performance


chr.info = read.table('hg19_genes.txt', sep='\t', header=T)
chr.info = chr.info[1:22,]

chr.info$info.content = chr.info$protein_coding/chr.info$length

# No correlation between information content (genes/length) and classification rate 
cor.test(mp$performance[1:22,'mean'], chr.info$info.content)


median(mp$performance$mean)
sd(mp$performance$mean)
mp$performance[17,'mean']

```

Individually none of these show very good classification rates, all hovering around 65%. However, more importantly is that Chr17 is only marginally better than any of the others (unless you exclude HGD) and all are below the 72% classification rate we get with all of the data



# Tumour Supressors

```{r ts, eval=F}

colnames(ccgenes)

tsgs = subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer))
tsgs = tsgs[grep('^X|Y',tsgs$Chr.Band, invert=T),]
nrow(tsgs)

loc = do.call(rbind.data.frame, strsplit(tsgs$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

tsgs = cbind(tsgs, loc)
tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol','Role.in.Cancer', 'Mutation.Types')], keep.extra.columns = T)
tsgs = sort(tsgs)

length(tsgs)

# average 1 per 9MB
avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)


ch = arrange(ch, -hazards)
features = makeGRangesFromDataFrame(ch, keep.extra.columns = T)

distanceTSG = data.frame(matrix(nrow=0,ncol=3))
for (i in 1:length(features)) {
  f = features[i]
  hits = nearest(f, tsgs, ignore.strand=T, select='all')
  if (length(hits) > 0) {
    distanceTSG = rbind(distanceTSG, cbind.data.frame(
      'features'=i,
      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
    )
  }
}

features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]

# In our dataset we should see 1 TSG/onc per 10MB
# But this is double the actual number we see
ncol(dysplasia.df)/round(avgTSGdist/5e6)







length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))



```


# Validation Dataset

20 Patients (10P, 10NP) set aside for validation purposes
`r pander(table(sum.validation.info$Status), caption='Validation cohort')`


```{r val_setup, echo=F, message=F, warning=F, include=F}
tileFile = paste('tile_patients_All_',as.character(tile.w),'.Rdata',sep='')
tile.files = sapply(tileFile, function(f) grep(f, analysis.files, value=T,fixed=T))

fauxlabel = rep(0, length(validation.patient.info$Samplename) )
names(fauxlabel) = validation.patient.info$Samplename
validation.df = set.up.data(tile.files, validation.patient.info$Samplename, fauxlabel)
dim(validation.df)

if (includeDemo) {
  ## Add in demographics
  validation.df = add.demo.tocv(validation.patient.info, validation.df)
}

```



```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}

pander(table(validation.patient.info[c('Status','p53.Status')]))

valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])

locs = do.call(rbind.data.frame, lapply(rownames(valCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

valCNdf = cbind.data.frame(locs, valCNdf)
valCNdf = melt(valCNdf, id.vars=colnames(locs))
valCNdf$chr = factor(valCNdf$chr, levels=c(1:22), ordered = T)

valCNdf$seg.type = '5MB'
valCNdf[with(valCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
valCNdf = merge(valCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

valCNdf = merge(valCNdf, validation.patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
valCNdf$Status = ifelse(valCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

valCNdf$cn = factor(sapply(valCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)


n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))
# Average cn value
valCNdf = valCNdf %>% group_by(Status, start, end) %>% mutate( 'mean.value' = value/length(unique(variable)) )
valCNdf$mean.value = valCNdf$value # but not going to use it 
cn.mtn.plot(subset(valCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))

head(valCNdf)

df = rbind(cbind(valCNdf, 'set'=rep('Val', nrow(valCNdf))), cbind(globalCNdf, 'set'=rep('Train', nrow(globalCNdf)) ) )
subset(df, seg.type != 'arms') %>% group_by(set,Status) %>% summarise( var(mean.value), mean(mean.value), max(mean.value), min(mean.value) )


```


```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
message(paste("loading", file))
load(file, verbose=T)

fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')

vpd = lapply(validation.patient.data, function(pt) {
  info = pt$info
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

samples = intersect(rownames(pred), validation.patient.info$Samplename)
validation.patient.info$Prediciton = NA
validation.patient.info[which(validation.patient.info$Samplename %in% samples), 'Prediction'] = pred[samples,]

for (pt in names(vpd)) 
  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]

back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back.v$brks = with(back.v, factor(brks, rev(levels(brks))))

back.v$Pathology = as.character(back.v$Pathology)
back.v$Pathology[which(back.v$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','IMC/HGD'), ordered = T)

```


```{r echo=F, fig.height=14,fig.width=10}
back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
back.v$labels = factor(back.v$Risk, levels=c('High','Moderate','Low'))
back.v$nl = factor(back.v$nl, ordered = T)

sq = c(0, seq(1, max(back.v$months.before.final), 12))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, 
                 labels=c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-length(sq)],
                  include.lowest = T, ordered_result = T)
back.v$brks = with(back.v, factor(brks, levels=rev(levels(brks))))


head(back.v)

m = melt(back.v %>% group_by(Patient, Status) %>% summarise( 'Low'=length(which(Risk == 'Low'))/length(Patient),
                                            'Moderate'=length(which(Risk == 'Moderate'))/length(Patient),
                                            'High'=length(which(Risk == 'High'))/length(Patient),
                                            'n.samples'=length(Patient)), id.vars=c('Patient','Status','n.samples'))


ggplot(m, aes(Patient,  value*100, group=variable, fill=variable)) + geom_bar(stat='identity', position=position_stack()) +
  geom_text(aes(label=n.samples, y=101)) + scale_fill_manual(values=c('green3','orange2','red2')) +
  facet_grid(~Status, scales='free_x') + labs(y='Percent of samples', title='Validation patient set')



grid.arrange(
  pred.tiles(subset(back.v, Status == 'P')) + labs(title='Validation Progressors') ,
  pred.tiles(subset(back.v, Status == 'NP')) + labs(title='Validation Non-Progressors'),
  ncol=2)

```


```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=8}
riskPath = rbind( cbind.data.frame( table(subset(back.v, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
                  cbind.data.frame( table(subset(back.v, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back.v %>% group_by(Status, Pathology) %>% summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 2))
riskPath$Risk = factor(riskPath$Risk, levels=c('High','Moderate','Low'), ordered = T)

p53Path = cbind.data.frame(table(subset(back.v, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back.v, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% summarise( 'nPath'=length(Pathology)), p53Path)
p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% summarise('nPath'=unique(nPath))

ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=c('red','orange2','green3')) + 
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5) +
  facet_grid(~Status) + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')

```

## Clinical Story

First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.  Likely a consequence of sample size. The labels show the proprotions of progressors that fell into each quintile in the training set.
```{r, echo=F, warning=F, message=F}

summary.v = table.path.risk(back.v)

pander(as.data.frame(summary.v))

write.table(summary.v, sep='\t', quote=F, row.names = F)
```

### Vignetts

```{r, echo=F, warning=F, message=F, fig.width=12, fig.height=4}
vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')

back$Risk = factor(back$Risk, levels=levels(back$labels), ordered = T)

vig.plot<-function(df, info) {
  scale = seq(min(info$Endoscopy.Year), max(info$Endoscopy.Year), 1)
  subtitle = with(info, paste(unique(Age.at.diagnosis),unique(Sex),': ', sep=''))
  if (!is.na(unique(info$Smoking)))                             
    subtitle = paste(subtitle, ifelse(unique(info$Smoking) == 'Y', 'smoker, ','non-smoker, '),sep='')

  subtitle = paste(subtitle, 'M', unique(info$Maximal), sep='')
    
  if (!is.na(unique(info$Circumference)))
    subtitle = paste(subtitle, 'C', info$Circumference, sep='')
  
  subtitle = paste(subtitle, " BE diagnosed ", unique(info$Initial.Endoscopy), sep='')

  ggplot(df, aes(Endoscopy.Year, nl)) + geom_tile(aes(fill=Risk), color='white') +
    scale_fill_manual(values=c('red3','orange', 'green3'), limits=levels(df$Risk),  name='Progression') +
    geom_text_repel(aes(label=Pathology), color='white', show.legend=F) +
    labs(y='Oesophageal Location',x='Months Prior to Endpoint', title='') +
    plot.theme + theme(axis.text.x=element_text(angle=45, hjust=1),legend.position='none') + 
    scale_x_continuous(breaks=scale) +
    labs(title=paste('Patient',unique(info$Patient)), subtitle=subtitle)
}

## P
vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34))
vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56))
vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20))
vig.plot(subset(back, Patient == 86), subset(patient.info, Patient == 86))


## NP
vig.plot(subset(back, Patient == 3), subset(patient.info, Patient == 3))
vig.plot(df=subset(back, Patient == 23),info= subset(patient.info, Patient == 23))
vig.plot(subset(back, Patient == 33), subset(patient.info, Patient == 33))
vig.plot(subset(back, Patient == 49), subset(patient.info, Patient == 49))
vig.plot(subset(back, Patient == 51), subset(patient.info, Patient == 51))
           


```



```{r, echo=F, warning=F, message=F}
ggplot(back.v, aes(Prediction)) + geom_histogram(aes(fill=..x..), bins=20, show.legend = F) +
  scale_fill_gradient2(high='red3',low='green',mid='orange', limits=c(0,1), midpoint=cutoff, name='P(P)') +
  geom_vline(xintercept=pred.confidence$r2, linetype='longdash') +
#  geom_label(data=(pred.confidence %>% rowwise %>% mutate( 'mid'=sum(r1,r2)/2 )), aes(x=mid, y=11, label=paste(round(perc,2),conf,sep=''),color=conf ), show.legend=F) + 
  scale_color_manual(values=c('grey','black')) + plot.theme + labs(title='Validation predictions', y='n Samples', x='Probability') 
```





## Clonal Heterogenity

```{r clonality, echo=F, message=F, warning=F, eval=F}
patient.info = subset(patient.info, Hospital.Research.ID != 'AH0222')

multi_samp = subset(patient.info, PID %in% names(which(table(patient.info$PID) > 1)))
pts = unique(multi_samp$Hospital.Research.ID)

## Variance in in CN per endoscopy
for (pt in pts) {

  pid = names(which(table(pg.samp[[pt]]$PID) > 1))

  for(i in 1:length(pid)) {  
    samples = subset(pg.samp[[pt]], PID == pid[i])$Samplename
    subset(pg.samp[[pt]], PID == pid[i])$Prediction
    
    x = apply( dysplasia.df[ samples,  ],2, var)
    summary(x)
    
    range(x)
    plot(x)
  }


  

}




```


## Cox model - Time to progress?

```{r cox, warning=F, message=F, echo=F, eval=F}
library(survival)

cox.pvalue<-function(cox, test='logrank') {
  pvalue = NULL  
  csum = summary(cox)
  
  if (test == 'logrank') {
    pvalue = csum$sctest[['pvalue']]  
  } else if (test == 'wald') {
    pvalue = csum$waldtest[['pvalue']]
  } else if (test == 'loglik') {
    pvalue = csum$logtest[['pvalue']]
  }
  return(pvalue)
}


prog.pred = lapply(pg.samp, function(df) {
  df = df[,c('Hospital.Research.ID', 'Status', 'Patient', 'PID', 'Status', 'Endoscopy.Year','Initial.Endoscopy', 'Final.Endoscopy', 'Pathology', 'Samplename','Prediction','Prediction.Dev.Resid')] %>% rowwise() %>% mutate(
      'Time.to.Final'=Final.Endoscopy-Endoscopy.Year)
  df = arrange(df, Endoscopy.Year, Pathology)

  df$Dup.PID = duplicated(df$PID)
  x = df[!df$Dup.PID,]
  
  for (i in 1:nrow(x)) {
    if (i == 1) {
      x[i,'tstart'] = 0
      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
      next
    }  

    x[i,'tstart'] = x[(i-1), 'tstop']

    # Final sample is different -- will leave it out later
    if (i == nrow(x)) {
      x[i, 'tstop'] = x[i,'tstart'] + 0.5
      break
    }

    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']

    # Samples taken in the same year
    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
      x[i,'tstop'] = x[i,'tstop'] + 0.5 
  }

  x$Diagnosis = as.integer(x$Pathology %in% c('HGD','IMC'))
  
  if (sum(x$Diagnosis) >= 1)
    x = x[-which(x$Diagnosis == 1),]

  if (nrow(x) > 0 && sum(x$Diagnosis) <= 0 && x$Status == 'P')
    x[nrow(x), 'Diagnosis'] = 1

  if (nrow(x) > 0) {
    x$Dup.Samples = ''
    for (j in 1:nrow(df[df$Dup.PID,]))  {
      row = df[df$Dup.PID,][j,]
      x[which(x$PID == row[['PID']]), 'Dup.Samples'] = paste(x[which(x$PID == row[['PID']]), 'Dup.Samples'], row[['Samplename']], sep=',')
    }
    x$Dup.Samples = sub('^,','',x$Dup.Samples)
  }

  return(x)
})
## By pathology?
pp = do.call(rbind, prog.pred)
pp[is.na(pp$tstop), 'tstop'] = 0.5
pp$Pathology = as.character(pp$Pathology)

sfit = survfit( Surv(tstart, tstop, Diagnosis, type='interval') ~ Pathology, data=pp, conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit,cens.shape=3,lty.est=1,back.white=F)
summary(sfit)

cox = coxph( Surv(tstart, tstop, Diagnosis) ~ Pathology, data=pp) 
ggsurv(survfit(cox),cens.shape=3,lty.est=1,back.white=F,
      main="coxph(Survival, Diagnosis) ~ Pathology" ) + 
      labs(subtitle=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(pp), 'samples'))
summary(cox)

## All genomic regions
sample.pred = merge( do.call(rbind.data.frame, prog.pred), dysplasia.df, by.x='Samplename', by.y='row.names' )
sample.pred$Pathology = as.character(sample.pred$Pathology)

gcols = grep('\\d+:', colnames(sample.pred), value=T)
sample.pred = sample.pred[c('tstart', 'tstop','Diagnosis','Pathology',gcols)]

# all
#cox = coxph( Surv(tstart, tstop, Diagnosis) ~., data=sample.pred)
#ggsurv(survfit(cox))
#summary(cox)

# glm selected features
cox2 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', 'Pathology',  ch$label)]) 
ggsurv(survfit(cox2))
summary(cox2)

# chr 17 only
cox17 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^17:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox17))
summary(cox17)


cox11 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^11:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox11))
summary(cox11)


cox13 = coxph( Surv(tstart, tstop, Diagnosis) ~ ., data=sample.pred[,c('tstart', 'tstop','Diagnosis', grep('^13:',colnames(sample.pred), value=T))]) 
ggsurv(survfit(cox13))
summary(cox13)


```


## CoxRFX

```{r echo=F, message=F, warning=F, eval=F}
plot.waldtest<-function(wt) {
  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
  if (length(grep('hazards', colnames(wt))) > 0) {
      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
        geom_point(aes(coef.x, coef.y, shape=gl, color=p.value<0.05, size=hazards), show.legend = F) +
        #scale_shape_manual(values=c(17,18), na.value=21) +
        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
        #geom_text_repel( aes( label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, p.value < 0.05 & !is.na(gl)), aes(coef.x,coef.y, label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
  } else {
    p = p + geom_point(alpha=0.5) + 
      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
  }
  return(p + labs(title='Wald test'))
}
```

```{r echo=T, message=F, warning=F, fig.height=8, fig.width=8, eval=F}

library(CoxHD)

coxDF = sample.pred

pathology = do.call(cbind.data.frame, lapply(unique(coxDF$Pathology), function(p) {
  cbind( as.integer(coxDF$Pathology == p))
}))
colnames(pathology) = unique(coxDF$Pathology)

## GLM regions only
coxDF = cbind(coxDF[,1:3],pathology[,c('BE','ID')], coxDF[,ch$label])
coxDF$Pathology = NULL
dim(coxDF)

gcols = colnames(coxDF)
cols = grep('^\\d+', colnames(coxDF))
colnames(coxDF)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(coxDF), value=T)), sep='')


surv = with(coxDF, Surv(tstart, tstop, Diagnosis))

groups = factor(c(rep('clinical',2),c(rep('genomic', length(grep('\\d+', colnames(coxDF)))))))
rfx = CoxRFX(coxDF[,-c(1:3)], surv, groups, verbose=F, nu=1)
summary(rfx)

plot(rfx)

wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
wt = merge(wt, ch[,c('label','hazards', 'coef','stability','gl')], by='label', all=T) 

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0


plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')

```









